\bf On semidefinite-representable sets over valuation fields

### 2025-05-15 article/paper_padic_spec.tex
SNF --> mathrm

We recall the definition of Smith Normal Form, in our context.
\begin{definition}[Smith Normal Form]\label{smith_nf}
  Let $M \in K^{m \times n}$ be of rank $r$. There exist unique integers $a_1 \leq \cdots \leq a_r$, and
  matrices $P \in \GL_n(\OK)$ and $Q \in \GL_m(\OK)$, such that
  $$
  SNF(M) := QMP^{-1} = \diag(\pi^{a_1}, \pi^{a_2}, \ldots, \pi^{a_r}, {\bf 0}_{\min(m,n)-r}).
  $$
  The matrix $SNF(M)$ is called the \emph{Smith Normal Form (SNF)} of $M$, and the elements $\pi^{a_i}$ the
  \emph{invariant factors}.
\end{definition}
\begin{remark}
  The valuation of the first invariant factor of $SNF(M)$ of a matrix
  $M \in K^{n \times n}$ is the minimum of the valuations of the entries of
  $M$. In particular, $SNF(M) \in \OK^{n \times n}$.
\end{remark}

@@

Thanks to Definition \ref{smith_nf}, (the matrix of) $f$
can be written $f=Q^{-1} \Delta P$
with $P \in \GL_n(\OK),$ $Q \in \GL_m(\OK)$
and $\Delta = SNF(f)\in \allmat_{m,n}(K)$ of the form $\diag(\lambda_1, \ldots, \lambda_r, {\bf 0}_{\min(m,n)-r})$
  where ${\bf 0}_{\min(m,n)-r}$ is a null square matrix of size $\min(m,n)-r$ for $\lambda_1,\dots,\lambda_r$ non-zero and $r \in \N.$

@@

\begin{proposition}\label{reducD}
  Let $A \in K^{d \times n}, b \in K^d, c\in K^n$, $D \in K^{m\times n}$,
  $e \in K^m$ and let $J \in K^{n \times R}$ be a matrix such that
  ${\rm Im}\, J = \ker D$ and $R = \dim \ker D$.
  Assume there exists $x_0 \in K^n$ such that $D x_0 = e$, and let
  $x \in \{z \in K^n : Dz = e\} = x_0 + \ker D$.
  Let $y \in K^R$ denote the unique element of $K^n$ such that
  $x = x_0 + J y$. Then:
  \begin{enumerate}
  \item\label{reducD_item1}
    $x$ is feasible for~\eqref{LP} with data $(A,b,c,D,e)$ if and only if 
    $y$ is feasible for~\eqref{LP} with data $(AJ,b+A x_0,J^T c,0,0)$.
  \item\label{reducD_item2}
    $x$ is a solution for~\eqref{LP} with data $(A,b,c,D,e)$ if and only if 
    $y$ is a solution for~\eqref{LP} with data $(AJ,b+A x_0,J^T c,0,0)$ and
    $\langle c,x^* \rangle = \langle J^T c,y^* \rangle$.
  \end{enumerate}
\end{proposition}

@@

The most interesting aspect of the SNF regarding the resolution of~\eqref{LP} is that the transition matrices are
invertible in $\OK$ and as such they preserve inequalities. %Indeed, for all $z \in K^n$ and $M \in \allmat_n(\OK)$ if $z \geq 0$, then $Mz \geq 0$.
%The latter property is a direct consequence of $\OK$ being a ring and the converse holds if the matrix $M$ is
%invertible in $\OK$. 
Therefore, for all matrix $M \in \allmat_{n,n}(K)$ with $S=SNF(M)$ and for all $z \in K^n$,
then $Mz \geq 0$ if and only if $Sz \geq 0$.

@@

\begin{proposition} \label{solsLP}
  Let $A \in \allmat_{d,n}(K), b \in K^d, c\in K^n$ and let $S = SNF(A) = QAP^{-1} \in \allmat_{d,n}(K)$, with
  transition matrices $P \in \GL_d(\OK)$ and $Q\in \GL_d(\OK)$. Define $b' = Qb$ and $c' = (P^{-1})^Tc$.
  Then
   \begin{enumerate}
   \item \label{solsLP_item1}
     $x \in K^n$ is feasible for \eqref{LP} with data $(A,b,c,0,0)$ if and only if $y = P x \in K^n$ is
     feasible for \eqref{LP} with data $(S,b',c',0,0)$.
   \item \label{solsLP_item2}
     $x^* \in K^n$ is a solution to \eqref{LP} with data $(A,b,c,0,0)$ if and only if $y^* = P x^*$ is
     a solution to \eqref{LP} with data $(S,b',c',0,0)$ and $\langle c,x^* \rangle = \langle c',y^* \rangle$.
   \end{enumerate}
\end{proposition}
\begin{proof}
  First of all, $Ax+b = AP^{-1}y +b \geq 0$ if and only if $Q(AP^{-1}y + b) = Sy + b' \geq 0$ because
  $Q \in GL_m(\OK)$, which yields \Cref{solsLP_item1}. Next, for all $c \in K^n$, we have $\langle c,
  x\rangle = \langle c, P^{-1}Px \rangle = \langle \left(P^{-1}\right)^T c, Px \rangle = \langle c',y\rangle$,
  which proves \Cref{solsLP_item2}.
\end{proof}

\begin{proposition}\label{prop:reduc}
  Let $(A,b,c), (S,b',c') \in \allmat_{d,n}(K) \times K^d \times K^n$ be as in \Cref{solsLP}, where
  $S=SNF(A)$. Then
  \begin{enumerate}
  \item \label{prop:reduc_it1}
    \eqref{LP} with data $(S,b',c',0,0)$ is feasible if and only if the $d-r$ last coefficients
    of $b'$ have nonnegative valuation, for $r = {\rm rank}\,S$.
  \item \label{prop:reduc_it2}
    Assume at least one of the $n-r$ last coefficients of $c'$ is non-zero. If \eqref{LP} with
    data $(S,b',c')$ is feasible, then it is unbounded and does not admit solutions.
  \end{enumerate}
\end{proposition}
\begin{proof}
  %%%%\corentin{préciser les notations}. 
  A vector $y = (y_1, \ldots, y_n)^T \in K^n$ is feasible for \eqref{LP} with data $(S,b',c',0)$
  if and only if $\val(s_i y_i + b_i) \geq 0$ for all $i = 1,\ldots,n$.
  However $s_i = 0$ for $n-r \leq i\leq n$. Therefore, if \eqref{LP} with data $(S,b',c',0)$ is feasible,
  necessarily the $d-r$ last coefficients of $b'$ have nonnegative valuation.
  Conversely, if the $d-r$ last coefficients of $b'$ are zero, the vector $y \in K^n$ such that
  $y_i = -\frac{-b'_i}{s_i}$ if $1 \leq i \leq r$, and $y_i =0$ otherwise, is feasible for \eqref{LP}
  with data $(S,b',c',0,0)$.

>>>>

We recall the definition of Smith Normal Form, in our context.
\begin{definition}[Smith Normal Form]\label{smith_nf}
  Let $M \in K^{m \times n}$ be of rank $r$. There exist unique integers $a_1 \leq \cdots \leq a_r$, and
  matrices $P \in \GL_n(\OK)$ and $Q \in \GL_m(\OK)$, such that
  $$
  \mathrm{SNF}(M) := QMP^{-1} = \diag(\pi^{a_1}, \pi^{a_2}, \ldots, \pi^{a_r}, {\bf 0}_{\min(m,n)-r}).
  $$
  The matrix $\mathrm{SNF}(M)$ is called the \emph{Smith Normal Form (SNF)} of $M$, and the elements $\pi^{a_i}$ the
  \emph{invariant factors}.
\end{definition}
\begin{remark}
  The valuation of the first invariant factor of $\mathrm{SNF}(M)$ of a matrix
  $M \in K^{n \times n}$ is the minimum of the valuations of the entries of
  $M$. In particular, $\mathrm{SNF}(M) \in \OK^{n \times n}$.
\end{remark}

@@

Thanks to Definition \ref{smith_nf}, (the matrix of) $f$
can be written $f=Q^{-1} \Delta P$
with $P \in \GL_n(\OK),$ $Q \in \GL_m(\OK)$
and $\Delta = \mathrm{SNF}(f)\in \allmat_{m,n}(K)$ of the form $\diag(\lambda_1, \ldots, \lambda_r, {\bf 0}_{\min(m,n)-r})$
  where ${\bf 0}_{\min(m,n)-r}$ is a null square matrix of size $\min(m,n)-r$ for $\lambda_1,\dots,\lambda_r$ non-zero and $r \in \N.$

@@

\begin{proposition}\label{reducD}
  Let $A \in K^{d \times n}, b \in K^d, c\in K^n$, $D \in K^{m\times n}$,
  $e \in K^m$ and let $J \in K^{n \times R}$ be a matrix such that
  ${\rm Im}\, J = \ker D$ and $R = \dim \ker D$.
  Assume there exists $x_0 \in K^n$ such that $D x_0 = e$, and let
  $x \in \{z \in K^n : Dz = e\} = x_0 + \ker D$.
  Let $y \in K^R$ denote the unique\footnote{\simone{pas sur que ce soit unique,
    la mappe $y\mapsto Jy$ est surjective sur $Im J=\ker D$, non pas injective, non ?
    Il y a un seul élément de $\ker D$, mais il peut y avoir plusieurs fibres.
    En tous cas, a-t-on a besoin de l'unicité dans la preuve ?}}
  element of $K^n$ such that
  $x = x_0 + J y$. Then:
  \begin{enumerate}
  \item\label{reducD_item1}
    $x$ is feasible for~\eqref{LP} with data $(A,b,c,D,e)$ if and only if 
    $y$ is feasible for~\eqref{LP} with data $(AJ,b+A x_0,J^T c,0,0)$.
  \item\label{reducD_item2}
    $x$ is a solution for~\eqref{LP} with data $(A,b,c,D,e)$ if and only if 
    $y$ is a solution for~\eqref{LP} with data $(AJ,b+A x_0,J^T c,0,0)$ and
    $\langle c,x^* \rangle = \langle J^T c,y^* \rangle$.
  \end{enumerate}
\end{proposition}

@@

Concerning the usage of SNF in the resolution of \eqref{LP}, note that the transition matrices are invertible in $\OK$ and as such they preserve inequalities.
%%%%Indeed, for all $z \in K^n$ and $M \in \allmat_n(\OK)$ if $z \geq 0$, then $Mz \geq 0$. The latter property is a direct consequence of $\OK$ being a ring and the converse holds if the matrix $M$ is invertible in $\OK$.
Therefore, for all matrix $M \in K^{n\times n}$ with $S = \mathrm{SNF}(M)$
and for all $z \in K^n$,
then $Mz \geq 0$ if and only if $Sz \geq 0$.

@@

\begin{proposition} \label{solsLP}
  Let $A \in \allmat_{d,n}(K), b \in K^d, c\in K^n$ and let $S = \mathrm{SNF}(A) = QAP^{-1} \in \allmat_{d,n}(K)$, with
  transition matrices $P \in \GL_d(\OK)$ and $Q\in \GL_d(\OK)$. Define $b' = Qb$ and $c' = (P^{-1})^Tc$.
  Then
   \begin{enumerate}
   \item \label{solsLP_item1}
     $x \in K^n$ is feasible for \eqref{LP} with data $(A,b,c,0,0)$ if and only if $y = P x \in K^n$ is
     feasible for \eqref{LP} with data $(S,b',c',0,0)$.
   \item \label{solsLP_item2}
     $x^* \in K^n$ is a solution to \eqref{LP} with data $(A,b,c,0,0)$ if and only if $y^* = P x^*$ is
     a solution to \eqref{LP} with data $(S,b',c',0,0)$ and $\langle c,x^* \rangle = \langle c',y^* \rangle$.
   \end{enumerate}
\end{proposition}
\begin{proof}
  First of all, $Ax+b = AP^{-1}y +b \geq 0$ if and only if $Q(AP^{-1}y + b) = Sy + b' \geq 0$ because
  $Q \in GL_m(\OK)$, which yields \Cref{solsLP_item1}. Next, for all $c \in K^n$, we have $\langle c,
  x\rangle = \langle c, P^{-1}Px \rangle = \langle \left(P^{-1}\right)^T c, Px \rangle = \langle c',y\rangle$,
  which proves \Cref{solsLP_item2}.
\end{proof}

\begin{proposition}\label{prop:reduc}
  Let $(A,b,c), (S,b',c') \in \allmat_{d,n}(K) \times K^d \times K^n$ be as in \Cref{solsLP}, where
  $S=\mathrm{SNF}(A)$. Then
  \begin{enumerate}
  \item \label{prop:reduc_it1}
    \eqref{LP} with data $(S,b',c',0,0)$ is feasible if and only if the $d-r$ last coefficients
    of $b'$ have nonnegative valuation, for $r = {\rm rank}\,S$.
  \item \label{prop:reduc_it2}
    Assume at least one of the $n-r$ last coefficients of $c'$ is non-zero. If \eqref{LP} with
    data $(S,b',c')$ is feasible, then it is unbounded and does not admit solutions.
  \end{enumerate}
\end{proposition}
\begin{proof}
  %%%%\corentin{préciser les notations}. 
  A vector $y = (y_1, \ldots, y_n)^T \in K^n$ is feasible for \eqref{LP} with data $(S,b',c',0)$
  if and only if $\val(s_i y_i + b_i) \geq 0$ for all $i = 1,\ldots,n$.
  However $s_i = 0$ for $n-r \leq i\leq n$. Therefore, if \eqref{LP} with data $(S,b',c',0)$ is feasible,
  necessarily the $d-r$ last coefficients of $b'$ have nonnegative valuation.
  Conversely, if the $d-r$ last coefficients of $b'$ are zero, the vector $y \in K^n$ such that
  $y_i = -\frac{-b'_i}{s_i}$ if $1 \leq i \leq r$, and $y_i =0$ otherwise, is feasible for \eqref{LP}
  with data $(S,b',c',0,0)$.
